News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2016-01-08 rupnik2015.tex
fixes after proofreading

The term event is vague and ambiguous, but for the practical purposes, we define it as ``any significant happening that is being reported about in the media.'' Examples of events would include shooting down of the Malaysia Airlines plane over Ukraine on July 18th, 2014 and HSBC's admittance of aiding their clients in tax evasion on February 9th, 2015 (Figure~\ref{fig:event2}). Events such as these are covered by many articles and the question is how to find all the articles in different languages that are describing a single event. Generally, events are more specific than general themes as the time component plays an important role -- for example, the two wars in Iraq would be considered as separate events.

@@

Based on the similarity function, we propose a novel algorithm for linking events/clusters across languages. The approach is based on learning a classification model from labelled data based on several sets of features. In addition to these features, cross-lingual similarity is also used to quickly identify a small list of potential linking candidates for each cluster.
This greatly increases the scalability of the system.

@@

Our goal is to build a system that monitors global media and analyzes how events are being reported on. Our approach consists of two steps: tracking events separately in each language (based on language detection and an online clustering approach) and then connecting them. The pipeline must be able to process millions of articles per day and perform billions of similarity computations each day. Both steps rely heavily on similarity computation, which implies that this must be highly scalable.

@@

We also used a term re-weighting scheme that adjusts for the fact that some words occur more frequently in general. A term weight should correspond to the importance of the term for the given corpus. The common weighting scheme is called \emph{Term Frequency Inverse Document Frequency} ($TFIDF$) weighting. An \emph{Inverse Document Frequency} ($IDF$) weight for the dictionary term $k$ is defined as $\log\left( \frac{N}{DF_k} \right)$, where $DF_k$ is the number of documents in the corpus which contain term $k$.
When building cross-lingual models, the IDF scores were computed with respect to the Wikipedia corpus. In the other part of our system, we computed TFIDF vectors on streams of news articles in multiple languages. There the IDF scores for each language changed dynamically - for each new document we computed the IDF of all news articles within a 10 day window.

@@

The original sum of squared correlations is formulated as an unconstrained problem:
\begin{equation*}
  \begin{aligned}
    & \underset{w_i \in \RR^{k}}{\text{maximize}}
    & & \sum_{i < j}^m  \rho(w_i^T Y_i, w_j^T Y_j)^2.
\end{aligned}
\end{equation*}
We solve a similar problem by restricting $i=1$ and omitting the optimization over non-hub language pairs.
Let $D_{i,i} \in \RR^{k \times k}$ denote the empirical covariance of $\mathcal{Y}_i$ and $D_{i,j}$ denote the empirical cross-covariance computed based on $\mathcal{Y}_i$ and $\mathcal{Y}_j$. We solve the following constrained (unit variance constraints) optimization problem:
\begin{equation}\label{squaredCorHubOriginal}
  \begin{aligned}
    & \underset{w_i \in \RR^{k}}{\text{maximize}}
    & & \sum_{i = 2}^m  \left(w_1^T D_{1,i} w_i \right)^2
    & \text{subject to}
    & & w_i^T D_{i,i} w_i = 1, \quad\forall i = 1,\ldots, m.
\end{aligned}
\end{equation}
The constraints $w_i^T D_{i,i} w_i$ can be simplified by using the Cholesky decomposition $D_{i,i} = K_i^T \cdot K_i$ and substitution: $y_i := K_i w_i$. By inverting the $K_i$ matrices and defining  $G_i := K_1^{-T} D_{1,i} K_i^{-1}$, the problem can be reformulated:
\begin{equation}\label{squaredCorHub}
  \begin{aligned}
    & \underset{y_i \in \RR^{k}}{\text{maximize}}
    & & \sum_{i = 2}^m  \left(y_1^T G_{i} y_i \right)^2
    & \text{subject to}
    & & y_i^T y_i = 1, \quad\forall i = 1,\ldots, m.
\end{aligned}
\end{equation}
A necessary condition for optimality is that the derivatives of the Lagrangian vanish. The Lagrangian of (\ref{squaredCorHub}) is expressed as:
$$  L(y_1, \ldots, y_m, \lambda_1, \ldots, \lambda_m) = \sum_{i = 2}^m  \left(y_1^T G_{i} y_i \right)^2 + \sum_{i=1}^m \lambda_i \left(y_i^T y_i - 1\right).$$
Stationarity conditions give us:
\begin{equation}\label{dLdx1}
 \frac{\partial}{\partial x_1} L = 0 \Rightarrow \sum_{i = 2}^m  \left(y_1^T G_{i} y_i \right) G_i y_i + \lambda_1 y_1 = 0,
\end{equation}
\begin{equation}\label{dLdxi}
\frac{\partial}{\partial x_i} L = 0 \Rightarrow \left(y_1^T G_{i} y_i \right) G_{i}^T y_1 + \lambda_i y_i = 0,~i > 1.
\end{equation}
Multiplying the equations (\ref{dLdxi}) with $y_i^T$ and applying the constraints, we can eliminate $\lambda_i$ which gives us:
\begin{equation}\label{eqy1yi}
G_{i}^T y_1 = \left(y_1^T G_{i} y_i \right) y_i,~i > 1.
\end{equation}
Plugging this into (\ref{dLdx1}), we obtain an eigenvalue problem:
$$\left( \sum_{i = 2}^m G_i G_{i}^T \right) y_1 + \lambda_1 y_1 = 0.$$
The eigenvectors of $\left( \sum_{i = 2}^m G_i G_{i}^T \right)$ solve the problem for the first language. The solutions for $y_i$ are obtained from (\ref{eqy1yi}): $y_i := \frac{G_{i}^T y_1}{\| G_{i}^T y_1 \|}$.
Note that the solution (\ref{squaredCorHubOriginal}) can be recovered by: $w_i := K_i^{-1} y_i$. The linear transformation of the $w$ variables are thus expressed as:
$$ Y_1 := \text{eigenvectors of} \sum_{i = 2}^m G_i G_{i}^T, $$
$$ W_1 = K_1^{-1} Y_1 $$
$$ W_i = K_i^{-1} G_{i}^T Y_1 N,$$
where $N$ is a diagonal matrix that normalizes $G_{i}^T Y_1$, with $N(j,j) := \frac{1}{\|G(_{i} Y_1(:,j)\|}$.

@@

The main application on which we test the cross-lingual similarity is cross-lingual event linking. In online media streams -- particularly news articles -- there is often duplication of reporting, different viewpoints or opinions, all centering around a single event. The same events are covered by many articles and the question we address is how to find all the articles in different languages that are describing a single event. In this paper we consider the problem of matching events from different languages. We do not address the problem of detection of events and instead base our evaluation on an online system for detection of world events, Event Registry. The events are represented by clusters of articles and so ultimately our problem reduces to finding suitable matchings between clusters with articles in different languages.

@@

The problem of cross-lingual event linking is to match monolingual clusters of news articles that describing the same event across languages. For example, we want to match a cluster of Spanish news articles and a cluster of English news articles that both describe the same earthquake.

@@

We are also seeking for an algorithm which does not do exhaustive comparison of all clusters, since that can become prohibitively expensive when working in a real-time setting. More specifically, we wish to avoid testing cluster $c_i$ with all the clusters from all the other languages. Performing exhaustive comparison would result in $O(|C|^2)$ tests, where $|C|$ is the number of all clusters (over all languages), which is not feasible when the number of clusters is on the order of tens of thousands. We address this by testing only clusters that are connected with at least one $k$-nearest neighbor (marked as \emph{candidate clusters} in Figure~\ref{fig:clusters}).

@@

We will describe the main dataset for building cross-lingual models which is based on Wikipedia and then present two sets of experiments. The first set of experiments
establishes that the hub based approach can deal with language pairs where little or no training data is available. The second set of experiments compares the main approaches
that we presented on the task of mate retrieval and the task of event linking. Finally, we examine how different choices of features impact the event linking performance.

@@

To investigate the empirical performance of the low-rank approximations we will test the algorithms on a large-scale, real-world multilingual dataset that we extracted from Wikipedia by using inter-language links for alignment. This  results in a large number of weakly comparable documents in more than $200$ languages. Wikipedia is a large source of multilingual data that is especially important for the languages for which no translation tools, multilingual dictionaries as Eurovoc~\cite{eurovoc}, or strongly aligned multilingual corpora as Europarl~\cite{europarl} are available. Documents in different languages are related with so called 'inter-language' links that can be found on the left of the Wikipedia page. The Wikipedia is constantly growing. There are currently 12 Wikipedias with more than 1 million %$10^6$
 articles, $52$ with more than 100k %$10^5$
 articles, $129$ with more than 10k articles, and $236$ with more than $1,000$ articles.

@@

We created a manually labelled dataset in order to evaluate cross-lingual event linking using two human annotators. The annotators were provided with an interface listing the articles, their content from and top concepts for a pair of clusters and their task was to determine if the clusters were equivalent or not (i.e., discuss same event). To obtain a pair of clusters $(c_i, c_j)$ to annotate, we first randomly chose a cluster $c_i$, used Algorithm~\ref{cluster_merge_algo1} to compute a set of potentially equivalent clusters $C$ and randomly chose a cluster $c_j \in C$. The dataset provided by the annotators contains 808 examples, of which 402 are equivalent clusters pairs and 406 are not. Clusters in each learning example are either in English, Spanish or German. Although Event Registry imports articles in other languages as well, we restricted our experiments to these three languages. We chose only these three languages since they have very large number of articles and clusters per day which makes the cluster linking problem hard due to large number of possible links.

@@

% cluster size?
We also analyzed how cluster size influences the accuracy of cluster linking. We would expect that if the tested pair of clusters has a larger number of articles then the classifier should be able to more accurately predict whether the clusters should be linked or not. The reasoning is that the large clusters would provide more document linking information (more articles mean more links to other similar articles) as well as more accurately aggregated semantic information. In the case of smaller clusters, the errors of the similarity models have greater impact which should decrease the performance of the classifier, too. To validate this hypothesis we have split the learning examples into two datasets -- one containing cluster pairs where the combined number of articles from both clusters is below 20 and one dataset where the combined number is 20 or more. The results of the experiment can be seen in Table~\ref{table:linkingEvalAlgosLargeSmall}. As it can be seen, the results confirm our expectations: for smaller clusters it is indeed harder to correctly predict if the cluster pair should be merged or not.

@@

\subsection{Remarks on the Reproducibility of Experiments}
We have made  both the code and data that were used in the experiments publicly available at \url{https://github.com/rupnikj/jair_paper.git}.
The manually labelled dataset used in the evaluation of event linking is available at in the ``dataset`` subfolder of the github repository. The included archive contains two folders: ``positive'' and ``negative'', where the first folder includes examples of cluster pairs in two languages that represent the same event and the second folder contains pairs of clusters in two languages that do not represent different events. Each example is a JSON file that contains at the top level information about a pair of clusters (including text of the articles) as well as a set of ``meta'' attributes, that correspond to features described in Section \ref{algo:features}.

>>>>

The term event is vague and ambiguous, but for the practical purposes, we define it as ``any significant happening that is being reported about in the media.'' Examples of events would include shooting down of the Malaysia Airlines plane over Ukraine on July 18th, 2014 (see Figure~\ref{fig:event2}) and HSBC's admittance of aiding their clients in tax evasion on February 9th, 2015. Events such as these are covered by many articles and the question is how to find all the articles in different languages that are describing a single event. Generally, events are more specific than general themes as the time component plays an important role -- for example, the two wars in Iraq would be considered as separate events.

@@

Using the similarity function, we propose a novel algorithm for linking events/clusters across languages. We pose the task as a classification problem based on several sets of features. In addition to these features, cross-lingual similarity is also used to quickly identify a small list of potential linking candidates for each cluster.
This greatly increases the scalability of the system.

@@

Our goal is to build a system that monitors global media and analyzes how events are being reported on. Our approach consists of two steps: tracking events separately in each language (based on language detection and an online clustering approach) and then connecting them. The pipeline must be able to process millions of articles per day and perform billions of similarity computations each day. Both steps rely heavily on similarity computation, which should therefore be highly scalable.

@@

We also used a term re-weighting scheme that adjusts for the fact that some words occur more frequently in general. A term weight should correspond to the importance of the term for the given corpus. The common weighting scheme is called \emph{Term Frequency Inverse Document Frequency} ($TFIDF$) weighting. An \emph{Inverse Document Frequency} ($IDF$) weight for the dictionary term $k$ is defined as $\log\left( \frac{N}{DF_k} \right)$, where $DF_k$ is the number of documents in the corpus which contain term $k$ and $N$ is the total number of documents in the corpus.
When building cross-lingual models, the IDF scores were computed with respect to the Wikipedia corpus. In the other part of our system, we computed TFIDF vectors on streams of news articles in multiple languages. There the IDF scores for each language changed dynamically - for each new document we computed the IDF of all news articles within a 10 day window.

@@

The original sum of squared correlations is formulated as an unconstrained problem:
\begin{equation*}
  \begin{aligned}
    & \underset{w_i \in \RR^{k}}{\text{maximize}}
    & & \sum_{i < j}^m  \rho(w_i^T Y_i, w_j^T Y_j)^2.
\end{aligned}
\end{equation*}
We solve a similar problem by restricting $i=1$ and omitting the optimization over non-hub language pairs.
Let $D_{i,i} \in \RR^{k \times k}$ denote the empirical covariance of $\mathcal{Y}_i$ and $D_{i,j}$ denote the empirical cross-covariance computed based on $\mathcal{Y}_i$ and $\mathcal{Y}_j$. We solve the following constrained (unit variance constraints) optimization problem:
\begin{equation}\label{squaredCorHubOriginal}
  \begin{aligned}
    & \underset{w_i \in \RR^{k}}{\text{maximize}}
    & & \sum_{i = 2}^m  \left(w_1^T D_{1,i} w_i \right)^2
    & \text{subject to}
    & & w_i^T D_{i,i} w_i = 1, \quad\forall i = 1,\ldots, m.
\end{aligned}
\end{equation}
The constraints $w_i^T D_{i,i} w_i$ can be simplified by using the Cholesky decomposition $D_{i,i} = K_i^T \cdot K_i$ and substitution: $y_i := K_i w_i$. By inverting the $K_i$ matrices and defining  $G_i := K_1^{-T} D_{1,i} K_i^{-1}$, the problem can be reformulated:
\begin{equation}\label{squaredCorHub}
  \begin{aligned}
    & \underset{y_i \in \RR^{k}}{\text{maximize}}
    & & \sum_{i = 2}^m  \left(y_1^T G_{i} y_i \right)^2
    & \text{subject to}
    & & y_i^T y_i = 1, \quad\forall i = 1,\ldots, m.
\end{aligned}
\end{equation}
A necessary condition for optimality is that the derivatives of the Lagrangian vanish. The Lagrangian of (\ref{squaredCorHub}) is expressed as:
$$  L(y_1, \ldots, y_m, \lambda_1, \ldots, \lambda_m) = \sum_{i = 2}^m  \left(y_1^T G_{i} y_i \right)^2 + \sum_{i=1}^m \lambda_i \left(y_i^T y_i - 1\right).$$
Stationarity conditions give us:
\begin{equation}\label{dLdx1}
 \frac{\partial}{\partial y_1} L = 0 \Rightarrow \sum_{i = 2}^m  \left(y_1^T G_{i} y_i \right) G_i y_i + \lambda_1 y_1 = 0,
\end{equation}
\begin{equation}\label{dLdxi}
\frac{\partial}{\partial y_i} L = 0 \Rightarrow \left(y_1^T G_{i} y_i \right) G_{i}^T y_1 + \lambda_i y_i = 0,~i > 1.
\end{equation}
Multiplying the equations (\ref{dLdxi}) with $y_i^T$ and applying the constraints, we can eliminate $\lambda_i$ which gives us:
\begin{equation}\label{eqy1yi}
G_{i}^T y_1 = \left(y_1^T G_{i} y_i \right) y_i,~i > 1.
\end{equation}
Plugging this into (\ref{dLdx1}), we obtain an eigenvalue problem:
$$\left( \sum_{i = 2}^m G_i G_{i}^T \right) y_1 + \lambda_1 y_1 = 0.$$
The eigenvectors of $\left( \sum_{i = 2}^m G_i G_{i}^T \right)$ solve the problem for the first language. The solutions for $y_i$ are obtained from (\ref{eqy1yi}): $y_i := \frac{G_{i}^T y_1}{\| G_{i}^T y_1 \|}$.
Note that the solution (\ref{squaredCorHubOriginal}) can be recovered by: $w_i := K_i^{-1} y_i$. The linear transformation of the $w$ variables are thus expressed as:
$$ Y_1 := \text{eigenvectors of} \sum_{i = 2}^m G_i G_{i}^T, $$
$$ W_1 = K_1^{-1} Y_1 $$
$$ W_i = K_i^{-1} G_{i}^T Y_1 N,$$
where $N$ is a diagonal matrix that normalizes $G_{i}^T Y_1$, with $N(j,j) := \frac{1}{\|G(_{i} Y_1(:,j)\|}$.

@@

The main application on which we test the cross-lingual similarity is cross-lingual event linking. In online media streams -- particularly news articles -- there is often duplication of reporting, different viewpoints or opinions, all centering around a single event. The same events are covered by many articles and the question we address is how to find all the articles in different languages that are describing a single event. In this paper we consider the problem of matching events from different languages. We do not address the problem of detection of events but instead base our evaluation on an online system for detection of world events, Event Registry. The events are represented by clusters of articles and so ultimately our problem reduces to finding suitable matchings between clusters with articles in different languages.

@@

The problem of cross-lingual event linking is to match monolingual clusters of news articles that describe the same event across languages. For example, we want to match a cluster of Spanish news articles and a cluster of English news articles that both describe the same earthquake.

@@

We also seek an algorithm which does not do exhaustive comparison of all clusters, since that can become prohibitively expensive when working in a real-time setting. More specifically, we wish to avoid testing cluster $c_i$ with all the clusters from all the other languages. Performing exhaustive comparison would result in $O(|C|^2)$ tests, where $|C|$ is the number of all clusters (over all languages), which is not feasible when the number of clusters is on the order of tens of thousands. We address this by testing only clusters that are connected with at least one $k$-nearest neighbor (marked as \emph{candidate clusters} in Figure~\ref{fig:clusters}).

@@

We will describe the main dataset for building cross-lingual models which is based on Wikipedia and then present two sets of experiments. The first set of experiments
establishes that the hub based approach can deal with language pairs where little or no training data is available. The second set of experiments compares the main approaches
that we presented on the task of mate retrieval and the task of event linking. In the mate retrieval task we are given a test set of document pairs, where each pair consists of a document and its translation. Given a query document from the test set, the goal is to retrieve its translation in the other language, which is also referred to as its \emph{mate} document. Finally, we examine how different choices of features impact the event linking performance.

@@

To investigate the empirical performance of the low-rank approximations we will test the algorithms on a large-scale, real-world multilingual dataset that we extracted from Wikipedia by using inter-language links for alignment. This results in a large number of weakly comparable documents in more than $200$ languages. Wikipedia is a large source of multilingual data that is especially important for the languages for which no translation tools, multilingual dictionaries as Eurovoc~\cite{eurovoc}, or strongly aligned multilingual corpora as Europarl~\cite{europarl} are available. Documents in different languages are related with inter-language links that can be found on the left of the Wikipedia page. The Wikipedia is constantly growing. There are currently 12 Wikipedias with more than 1 million %$10^6$
 articles, $52$ with more than 100k %$10^5$
 articles, $129$ with more than 10k articles, and $236$ with more than $1,000$ articles.

@@

We created a manually labelled dataset in order to evaluate cross-lingual event linking using two human annotators. The annotators were provided with an interface listing the articles, their content and top concepts for a pair of clusters. Their task was to determine if the clusters were equivalent or not (i.e., discuss same event). To obtain a pair of clusters $(c_i, c_j)$ to annotate, we first randomly chose a cluster $c_i$, used Algorithm~\ref{cluster_merge_algo1} to compute a set of potentially equivalent clusters $C$ and randomly chose a cluster $c_j \in C$. The dataset provided by the annotators contains 808 examples, of which 402 are equivalent clusters pairs and 406 are not. Clusters in each learning example are either in English, Spanish or German. Although Event Registry imports articles in other languages as well, we restricted our experiments to these three languages. We chose only these three languages since they have very large number of articles and clusters per day which makes the cluster linking problem hard due to large number of possible links.

@@

% cluster size?
We also analyzed how cluster size influences the accuracy of cluster linking. Our intuition was that linking large cluster pairs is easier than linking clusters with few articles. The reasoning is that the large clusters would provide more document linking information (more articles mean more links to other similar articles) as well as more accurately aggregated semantic information. In the case of smaller clusters, the errors of the similarity models have greater impact which should decrease the performance of the classifier, too. To validate this hypothesis we have split the learning examples into two datasets -- one containing cluster pairs where the combined number of articles from both clusters is below 20 and one dataset where the combined number is 20 or more. The results of the experiment can be seen in Table~\ref{table:linkingEvalAlgosLargeSmall}. As it can be seen, the results confirm our expectations: for smaller clusters it is indeed harder to correctly predict if the cluster pair should be merged or not.

@@

\subsection{Remarks on the Reproducibility of Experiments}
We have made  both the code and data that were used in the experiments publicly available at \url{https://github.com/rupnikj/jair_paper.git}.
The manually labelled dataset used in the evaluation of event linking is available at in the ``dataset'' subfolder of the github repository. The included archive contains two folders: ``positive'' and ``negative'', where the first folder includes examples of cluster pairs in two languages that represent the same event and the second folder contains pairs of clusters in two languages that do not represent different events. Each example is a JSON file that contains at the top level information about a pair of clusters (including text of the articles) as well as a set of ``meta'' attributes, that correspond to features described in Section \ref{algo:features}.
